### 2014-01-16 databases/databases_chapter.tex
Update encyclopedia chapters to include required definitions and section headings

@@

First, there are databases of computational models and model parts.  
The BioModels database (\cite{BioModels2010,BioModels2013}, \url{http://www.ebi.ac.uk/biomodels-main}) is a general repository of computational models of biological processes.  
The SenseLab database, ModelDB, (\cite{migliore_modeldb_2003}, \url{http://senselab.med.yale.edu/modeldb}) compiles user-submitted computational models developed specifically for simulating the electrophysiological and neurochemical properties of single neurons and networks of neurons (e.g., multi-compartment Hodgkin-Huxley type conductance based models \cite{hodgkin_quantitative_1952}). 
The Physiome Model Repository (\cite{physiomemodelrepository_2011}, \url{http://www.cellml.org/tools/pmr} is a software suite that facilitates storage and management of models, focusing on those described in CellML, a standard markup language for biological models. 
Similarly, OpenSourceBrain (\cite{OSB_2012}, \url{http://www.opensourcebrain.org}) is a community platform for collaborative development of computational neuron and network models that utilizes open standards such as NeuroML (\cite{gleeson_neuroml:_2010}) to facilitate interoperability and visualization of neuroscience models developed by different researchers.

Second, there are databases containing structured information specific to neurons and their properties. 
For example, information on the detailed shapes of neurons (i.e., their morphology) is being compiled by NeuroMorpho (\cite{neuromorpho_2007}, \url{http://neuromorpho.org}) which contains user-submitted neuron morphological reconstructions made using the NeuroLucida format \cite{glaser_neuron_1990}. 
Measured electrophysiological properties of neuron types, and the metadata associated with these measurements, is catalogued in the NeuroElectro Project (\cite{neuroelectro_2013}, \url{http://neuroelectro.org}).  
Similarly, other SenseLab databases, including NeuronDB and CellPropDB (\cite{crasto_senselab:_2007}, \url{http://senselab.med.yale.edu/NeuronDB}), contain information on the ionic currents and neurotransmitters expressed by each neuron and how these are distributed with respect to neuronal morphology. 
Detailed information on ion channel subtypes, including voltage and temporal dynamics, genetic homology, and corresponding literature references is being compiled by Channelpedia (\cite{ranjan_channelpedia:_2011}, \url{http://channelpedia.net}), a subproject within the Blue Brain Project \citep{markram_blue_2006}.

Third, there are databases focusing on the anatomical organization of the brain.  
BrainInfo (\url{http://braininfo.rprc.washington.edu}) provides general information about brain areas, including what they do, where they are located, and what they contain.  
The Allen Institute for Brain Sciences provides brain-wide gene expression atlases, where the expression of each of the genes in the mammalian genome has been systematically quantified throughout the brain for a number of animal species and across stages of neural development (\cite{lein_genome-wide_2007}, \url{http://brain-map.org}). 
Similarly, the Allen Institute also provides information on the anatomical connectivity of different brain regions.
Parallel to this effort is The Human Connectome Project (\cite{humanconnectomeproject_2013}, \url{http://www.humanconnectomeproject.org}), a large-scale effort to map complete structural and functional neural connections \textit{in vivo} in individual humans. 
BrainMap (\cite{brainmap_2005}, \url{http://www.brainmap.org}) consists of a database and related software to search published functional and structural human neuroimaging experiments.  
CoCoMac (\cite{cocomac_2001,cocomac_2012}, \url{http://cocomac.g-node.org}), in contrast, is focused on the primate brain, containing records of tracing studies in the macaque. 
The Brain Architecture Management System (BAMS, \cite{bota_brain_2005}, \url{http://brancusi1.usc.edu}) contains neural connectivity information across species that has been manually curated from the existing research literature.  
Finally, Cell Centered Database (\cite{cellcentered_2003,cellcentered_2009}, \url{http://ccdb.ucsd.edu}) focuses on images from light and electron microscopy, ranging from whole brain areas to subcellular compartments.

In addition to these neuroscience domain-specific databases are uber-databases that provide linking facilities for cross-resource data integration. 
For example, NeuroLex (\cite{larson_neurolex.org:_2013}, \url{http://neurolex.org}), provides a platform for community annotation of neuron types on the basis of morphological, neurochemical, or electrophysiological properties. 
Given this wealth of neuroscience resources, the Neuroscience Information Framework (NIF, \cite{gardner_neuroscience_2008},\url{http://www.neuinfo.org}), provides tools for semantically searching across these diverse databases through the development and incorporation of neuroscience domain-specific ontologies \cite{bug_nifstd_2008,larson_ontologies_2009,hamilton_ontological_2012,imam_development_2012}. 
For example, in NIF, the search query ``mitral cell" returns a number of database records including relevant research literature from PubMed, computational models from ModelDB, and connectivity information from BAMS.

@@

\bibliographystyle{plainnat}
\nobibliography{shreejoy}

>>>>

\section*{Neuroscience Databases}

\subsection*{Synonyms}
Data repositories

\subsection*{Definition}
Neuroscience databases are curated collections of data from neuroscience experiments that can be used to test models or compare with the results of novel experiments.  They are often but not always relational, and are frequently free and public to maximize their utility to other researchers.

\subsection*{Detailed Description}

@@

\subsubsection*{Computational Models}
First, there are databases of computational models and model parts.  
The BioModels database (\cite{BioModels2010,BioModels2013}, \url{http://www.ebi.ac.uk/biomodels-main}) is a general repository of computational models of biological processes.  
The SenseLab database, ModelDB, (\cite{migliore_modeldb_2003}, \url{http://senselab.med.yale.edu/modeldb}) compiles user-submitted computational models developed specifically for simulating the electrophysiological and neurochemical properties of single neurons and networks of neurons (e.g., multi-compartment Hodgkin-Huxley type conductance based models \cite{hodgkin_quantitative_1952}). 
The Physiome Model Repository (\cite{physiomemodelrepository_2011}, \url{http://www.cellml.org/tools/pmr} is a software suite that facilitates storage and management of models, focusing on those described in CellML, a standard markup language for biological models. 
Similarly, OpenSourceBrain (\cite{OSB_2012}, \url{http://www.opensourcebrain.org}) is a community platform for collaborative development of computational neuron and network models that utilizes open standards such as NeuroML (\cite{gleeson_neuroml:_2010}) to facilitate interoperability and visualization of neuroscience models developed by different researchers.

\subsubsection*{Neuron Properties}
Second, there are databases containing structured information specific to neurons and their properties. 
For example, information on the detailed shapes of neurons (i.e., their morphology) is being compiled by NeuroMorpho (\cite{neuromorpho_2007}, \url{http://neuromorpho.org}) which contains user-submitted neuron morphological reconstructions made using the NeuroLucida format \cite{glaser_neuron_1990}. 
Measured electrophysiological properties of neuron types, and the metadata associated with these measurements, is catalogued in the NeuroElectro Project (\cite{neuroelectro_2013}, \url{http://neuroelectro.org}).  
Similarly, other SenseLab databases, including NeuronDB and CellPropDB (\cite{crasto_senselab:_2007}, \url{http://senselab.med.yale.edu/NeuronDB}), contain information on the ionic currents and neurotransmitters expressed by each neuron and how these are distributed with respect to neuronal morphology. 
Detailed information on ion channel subtypes, including voltage and temporal dynamics, genetic homology, and corresponding literature references is being compiled by Channelpedia (\cite{ranjan_channelpedia:_2011}, \url{http://channelpedia.net}), a subproject within the Blue Brain Project \citep{markram_blue_2006}.

\subsubsection*{Anatomical Reference}
Third, there are databases focusing on the anatomical organization of the brain.  
BrainInfo (\url{http://braininfo.rprc.washington.edu}) provides general information about brain areas, including what they do, where they are located, and what they contain.  
The Allen Institute for Brain Sciences provides brain-wide gene expression atlases, where the expression of each of the genes in the mammalian genome has been systematically quantified throughout the brain for a number of animal species and across stages of neural development (\cite{lein_genome-wide_2007}, \url{http://brain-map.org}). 
Similarly, the Allen Institute also provides information on the anatomical connectivity of different brain regions.
Parallel to this effort is The Human Connectome Project (\cite{humanconnectomeproject_2013}, \url{http://www.humanconnectomeproject.org}), a large-scale effort to map complete structural and functional neural connections \textit{in vivo} in individual humans. 
BrainMap (\cite{brainmap_2005}, \url{http://www.brainmap.org}) consists of a database and related software to search published functional and structural human neuroimaging experiments.  
CoCoMac (\cite{cocomac_2001,cocomac_2012}, \url{http://cocomac.g-node.org}), in contrast, is focused on the primate brain, containing records of tracing studies in the macaque. 
The Brain Architecture Management System (BAMS, \cite{bota_brain_2005}, \url{http://brancusi1.usc.edu}) contains neural connectivity information across species that has been manually curated from the existing research literature.  
Finally, Cell Centered Database (\cite{cellcentered_2003,cellcentered_2009}, \url{http://ccdb.ucsd.edu}) focuses on images from light and electron microscopy, ranging from whole brain areas to subcellular compartments.

\subsubsection*{Integration}
In addition to these neuroscience domain-specific databases are uber-databases that provide linking facilities for cross-resource data integration. 
For example, NeuroLex (\cite{larson_neurolex.org:_2013}, \url{http://neurolex.org}), provides a platform for community annotation of neuron types on the basis of morphological, neurochemical, or electrophysiological properties. 
Given this wealth of neuroscience resources, the Neuroscience Information Framework (NIF, \cite{gardner_neuroscience_2008},\url{http://www.neuinfo.org}), provides tools for semantically searching across these diverse databases through the development and incorporation of neuroscience domain-specific ontologies \cite{bug_nifstd_2008,larson_ontologies_2009,hamilton_ontological_2012,imam_development_2012}. 
For example, in NIF, the search query ``mitral cell" returns a number of database records including relevant research literature from PubMed, computational models from ModelDB, and connectivity information from BAMS.

@@

\subsection*{Cross-References/Related Terms}
None

\bibliographystyle{abbrv}
\bibliography{shreejoy}

### 2014-01-16 databases/neuroelectro_chapter.tex
Update encyclopedia chapters to include required definitions and section headings

Computational neuroscientists have argued that the electrophysiological properties of neurons determine their computational roles.
Furthermore, models containing or concerning neurons can be constrained or validated in part using empirical data on single cell electrophysiology.
However, compared with imaging or genetic data, these data are infrequently published in a format that permits them to be easily discovered in the literature.
The NeuroElectro Project (\url{http://neuroelectro.org}, \cite{neuroelectro_2013}) solves this problem by providing a database on the electrophysiological diversity of mammalian neuron types.

This database is populated primarily using automated text extraction from tables of journal articles using Python.
A subset of records have been manually validated against the primary literature.
Neuron types are identified using the listing of neuron types provided by NeuroLex (\cite{larson_neurolex.org:_2013,hamilton_ontological_2012}, \url{http://neurolex.org});
Electrophysiological property types are identified using a custom ontology which formalizes existing community standards \cite{ascoli_petilla_2008}.

While NeuroElectro contains information about 27 distinct electrophysiological properties, those most heavily represented include resting membrane potential, input resistance, action potential height and width, etc.
The list of journals containing the source information include those from the Elsevier, Highwire, Wiley, and Oxford family of journals, including The Journal of Neuroscience, The Journal of Physiology, and The Journal of Neurophysiology, among others.
NeuroElectro also contains details on each study’s experimental conditions (like species or electrode type used), the so-called ``metadata'', obtained using similar text extraction techniques applied to article methods sections.

The NeuroElectro web interface (\url{http://neuroelectro.org} also invites users to contribute to the database’s content.
For example, users can suggest relevant articles on a specific neuron type as well as validate content that has been extracted via the automated algorithms.
Furthermore, users can elect to become ``content-experts” on a specific neuron type and can then modify the database’s content directly.

NeuroElectro also contains analyses of the resulting database yielding novel relationships on the electrophysiological similarity of different neuron types throughout the brain.
Among these are hierarchical classifications of neuron types based on similarity of electrophysiological properties.
The project maintainers are also aiming to relate gene expression patterns in specific cell types to those cells' electrophysiological properties, using expression data obtained through the Allen Brain Atlas (\cite{lein_genome-wide_2007}, \url{http://brain-map.org}).

\bibliographystyle{plainnat}
\nobibliography{shreejoy}
\end{document}

>>>>

\section*{The NeuroElectro Project}

\subsection*{Synonyms}
neuroelectro, neuroelectro.org

\subsection*{Definition}
The NeuroElectro Project is an effort to curate data about the electrophysiological properties
of neurons.  It contains experimental data extracted from the literature about dozens of
basic properties (such as resting membrane potential, action potential width, etc.) for hundreds of distinct types of neurons.  It can be browsed online at http://neuroelectro.org.

\subsection*{Detailed Description}
Computational neuroscientists have argued that the electrophysiological properties of neurons determine their computational roles.
Furthermore, models containing or concerning neurons can be constrained or validated in part using empirical data on single cell electrophysiology.
However, compared with imaging or genetic data, these data are infrequently published in a format that permits them to be easily discovered in the literature.
The NeuroElectro Project (\url{http://neuroelectro.org}, \cite{neuroelectro_2013}) solves this problem by providing a database on the electrophysiological diversity of mammalian neuron types.

\subsubsection*{Implementation}
This database is populated primarily using automated text extraction from tables of journal articles using Python.
A subset of records have been manually validated against the primary literature.
Neuron types are identified using the listing of neuron types provided by NeuroLex (\cite{larson_neurolex.org:_2013,hamilton_ontological_2012}, \url{http://neurolex.org});
Electrophysiological property types are identified using a custom ontology which formalizes existing community standards \cite{ascoli_petilla_2008}.

\subsubsection*{Data provenance}
While NeuroElectro contains information about 27 distinct electrophysiological properties, those most heavily represented include resting membrane potential, input resistance, action potential height and width, etc.
The list of journals containing the source information include those from the Elsevier, Highwire, Wiley, and Oxford family of journals, including The Journal of Neuroscience, The Journal of Physiology, and The Journal of Neurophysiology, among others.
NeuroElectro also contains details on each study’s experimental conditions (like species or electrode type used), the so-called ``metadata'', obtained using similar text extraction techniques applied to article methods sections.

\subsubsection*{Participation}
The NeuroElectro web interface (\url{http://neuroelectro.org} also invites users to contribute to the database’s content.
For example, users can suggest relevant articles on a specific neuron type as well as validate content that has been extracted via the automated algorithms.
Furthermore, users can elect to become ``content-experts" on a specific neuron type and can then modify the database’s content directly.

\subsubsection*{Scientific Goals}
NeuroElectro also contains analyses of the resulting database yielding novel relationships on the electrophysiological similarity of different neuron types throughout the brain.
Among these are hierarchical classifications of neuron types based on similarity of electrophysiological properties.
The project maintainers are also aiming to relate gene expression patterns in specific cell types to those cells' electrophysiological properties, using expression data obtained through the Allen Brain Atlas (\cite{lein_genome-wide_2007}, \url{http://brain-map.org}).

\subsection*{Cross-References/Related Terms}
None

\bibliographystyle{abbrv}
\bibliography{shreejoy}
\end{document}
